# RHC data description

```{r setup01, include=FALSE}
require(tableone)
require(Publish)
require(MatchIt)
require(cobalt)
```

There is a widespread belief among cardiologists that the right heart catheterization (RHC hereafter; a monitoring device for measurement of cardiac function) is helpful in managing critically ill patients in the intensive care unit. @connors1996effectiveness examined the association of 

- *RHC use* during the first 24 hours of care in the intensive care unit and 
- a number of health-outcomes such as *length of stay* (hospital).

## Data download

```{block, type='rmdcomment'}
Data is freely available from [Vanderbilt Biostatistics](https://hbiostat.org/data/).
```


```{r, cache=TRUE}
# load the dataset
ObsData <- read.csv("https://hbiostat.org/data/repo/rhc.csv", header = TRUE)
saveRDS(ObsData, file = "data/rhc.RDS")
```

## Analytic data

Below we show the process of creating the analytic data (optional).

```{r, warning=FALSE}
# add column for outcome Y: length of stay 
# Y = date of discharge - study admission date
# Y = date of death - study admission date if date of discharge not available
ObsData$Y <- ObsData$dschdte - ObsData$sadmdte
ObsData$Y[is.na(ObsData$Y)] <- ObsData$dthdte[is.na(ObsData$Y)] - 
  ObsData$sadmdte[is.na(ObsData$Y)]
# remove outcomes we are not examining in this example
ObsData <- dplyr::select(ObsData, 
                         !c(dthdte, lstctdte, dschdte, death, t3d30, dth30, surv2md1))
# remove unnecessary and problematic variables 
ObsData <- dplyr::select(ObsData, 
                         !c(sadmdte, ptid, X, adld3p, urin1, cat2))

# convert all categorical variables to factors 
factors <- c("cat1", "ca", "cardiohx", "chfhx", "dementhx", "psychhx", 
             "chrpulhx", "renalhx", "liverhx", "gibledhx", "malighx", 
             "immunhx", "transhx", "amihx", "sex", "dnr1", "ninsclas", 
             "resp", "card", "neuro", "gastr", "renal", "meta", "hema", 
             "seps", "trauma", "ortho", "race", "income")
ObsData[factors] <- lapply(ObsData[factors], as.factor)
# convert our treatment A (RHC vs. No RHC) to a binary variable
ObsData$A <- ifelse(ObsData$swang1 == "RHC", 1, 0)
ObsData <- dplyr::select(ObsData, !swang1)
# Categorize the variables to match with the original paper
ObsData$age <- cut(ObsData$age,breaks=c(-Inf, 50, 60, 70, 80, Inf),right=FALSE)
ObsData$race <- factor(ObsData$race, levels=c("white","black","other"))
ObsData$sex <- as.factor(ObsData$sex)
ObsData$sex <- relevel(ObsData$sex, ref = "Male")
ObsData$cat1 <- as.factor(ObsData$cat1)
levels(ObsData$cat1) <- c("ARF","CHF","Other","Other","Other",
                          "Other","Other","MOSF","MOSF")
ObsData$ca <- as.factor(ObsData$ca)
levels(ObsData$ca) <- c("Metastatic","None","Localized (Yes)")
ObsData$ca <- factor(ObsData$ca, levels=c("None",
                                          "Localized (Yes)","Metastatic"))
# Rename variables
names(ObsData) <- c("Disease.category", "Cancer", "Cardiovascular", 
                    "Congestive.HF", "Dementia", "Psychiatric", "Pulmonary", 
                    "Renal", "Hepatic", "GI.Bleed", "Tumor", 
                    "Immunosupperssion", "Transfer.hx", "MI", "age", "sex", 
                    "edu", "DASIndex", "APACHE.score", "Glasgow.Coma.Score", 
                    "blood.pressure", "WBC", "Heart.rate", "Respiratory.rate", 
                    "Temperature", "PaO2vs.FIO2", "Albumin", "Hematocrit", 
                    "Bilirubin", "Creatinine", "Sodium", "Potassium", "PaCo2", 
                    "PH", "Weight", "DNR.status", "Medical.insurance", 
                    "Respiratory.Diag", "Cardiovascular.Diag", 
                    "Neurological.Diag", "Gastrointestinal.Diag", "Renal.Diag",
                    "Metabolic.Diag", "Hematologic.Diag", "Sepsis.Diag", 
                    "Trauma.Diag", "Orthopedic.Diag", "race", "income", 
                    "Y", "A")
saveRDS(ObsData, file = "data/rhcAnalytic.RDS")
```

## Notations

|Notations| Example in RHC study|
|---|---|
|$A$: Exposure status  | RHC |  
|$Y$: Observed outcome  | length of stay  |  
|$L$: Covariates  | See below |  

## Variables

```{r vars, cache=TRUE, echo = TRUE}
baselinevars <- names(dplyr::select(ObsData, 
                         !c(A,Y)))
baselinevars
```

## Table 1 stratified by RHC exposure

```{block, type='rmdcomment'}
Only for some demographic and co-morbidity variables; match with Table 1 in @connors1996effectiveness.
```


```{r tab0, cache=TRUE, echo = TRUE}
require(tableone)
tab0 <- CreateTableOne(vars = c("age", "sex", "race", "Disease.category", "Cancer"),
                       data = ObsData, 
                       strata = "A", 
                       test = FALSE)
print(tab0, showAllLevels = FALSE, )
```

```{block, type='rmdcomment'}
Only outcome variable (Length of stay); slightly different than Table 2 in @connors1996effectiveness (means 20.5 vs. 25.7; and medians 13 vs. 17).
```

```{r tab1, cache=TRUE, echo = TRUE}
tab1 <- CreateTableOne(vars = c("Y"),
                       data = ObsData, 
                       strata = "A", 
                       test = FALSE)
print(tab1, showAllLevels = FALSE, )
median(ObsData$Y[ObsData$A==0]); median(ObsData$Y[ObsData$A==1])
```

## Basic regression analysis

### Crude analysis

```{r reg1, cache=TRUE, echo = TRUE, results='hide'}
# adjust the exposure variable (primary interest)
fit0 <- lm(Y~A, data = ObsData)
require(Publish)
crude.fit <- publish(fit0, digits=1)$regressionTable[2,]
```

```{r reg1c, cache=TRUE, echo = TRUE}
crude.fit
```


### Adjusted analysis

```{r reg2, cache=TRUE, echo = TRUE, results='hide'}
# adjust the exposure variable (primary interest) + covariates
out.formula <- as.formula(paste("Y~ A +", 
                               paste(baselinevars, 
                                     collapse = "+")))
fit1 <- lm(out.formula, data = ObsData)
adj.fit <- publish(fit1, digits=1)$regressionTable[2,]
```

```{r, cache=TRUE, echo = TRUE}
saveRDS(fit1, file = "data/adjreg.RDS")
```

```{r reg2a, cache=TRUE, echo = TRUE}
out.formula
adj.fit
```

### Regression diagnostics

```{r reg2a578, cache=TRUE, echo = TRUE}
plot(fit1)
```

```{block, type='rmdcomment'}
Diagnostics do not necessarily look so good. 
```


## Comparison with literature

```{block, type='rmdcomment'}
@connors1996effectiveness conducted a propensity score matching analysis. 
```

Table 5 in @connors1996effectiveness showed that, after propensity score pair (1-to-1) matching, means of length of stay ($Y$), when stratified by RHC ($A$) were not significantly different ($p = 0.14$). 

### PSM in RHC data

We also conduct propensity score pair matching analysis, as follows. 

```{block, type='rmdcomment'}
**Note**: In this workshop, we will not cover Propensity Score Matching (PSM). If you want to learn more about this, feel free to check out this other workshop: [Understanding Propensity Score Matching](https://ehsanx.github.io/psw/) and the [video recording](https://www.youtube.com/watch?v=u4Nl7gnDEAY) on youtube.
```

```{r ps16854, cache=TRUE, echo = TRUE}
set.seed(111)
require(MatchIt)
ps.formula <- as.formula(paste("A~", 
                paste(baselinevars, collapse = "+")))
PS.fit <- glm(ps.formula,family="binomial", 
              data=ObsData)
ObsData$PS <- predict(PS.fit, 
                      newdata = ObsData, type="response") 
```



```{r ps2, cache=TRUE, echo = TRUE}
logitPS <-  -log(1/ObsData$PS - 1)  
match.obj <- matchit(ps.formula, data =ObsData,
                     distance = ObsData$PS,
                     method = "nearest", replace=FALSE,
                     ratio = 1,
                     caliper = .2*sd(logitPS))
```

#### PSM diagnostics

```{r ps2x, cache=TRUE, echo = TRUE}
require(cobalt)
bal.plot(match.obj,  
         var.name = "distance", 
         which = "both", 
         type = "histogram",  
         mirror = TRUE)
bal.tab(match.obj, un = TRUE, 
        thresholds = c(m = .1))
```


```{r ps2b, cache=TRUE, echo = TRUE, fig.height=10, fig.width=5}
love.plot(match.obj, binary = "std", 
          thresholds = c(m = .1))  
```

The love plot suggests satisfactory propensity score matching (all SMD < 0.1).

#### PSM results

##### p-value

```{r ps3, cache=TRUE, echo = TRUE}
matched.data <- match.data(match.obj)   
tab1y <- CreateTableOne(vars = c("Y"),
               data = matched.data, strata = "A", 
               test = TRUE)
print(tab1y, showAllLevels = FALSE, 
      test = TRUE)
```

```{block, type='rmdcomment'}
Our conclusion based on propensity score pair matched data ($p  \lt 0.001$) is different than Table 5 in @connors1996effectiveness ($p = 0.14$). Variability in results for 1-to-1 matching is possible, and modelling choices may be different (we used caliper option here).
```

##### Treatment effect

- We can also estimate the effect of `RHC` on `length of stay` using propensity score-matched sample:

```{r ps12ryy, cache=TRUE, echo = TRUE}
fit.matched <- glm(Y~A,
            family=gaussian,  
            data = matched.data)  
publish(fit.matched)
```

```{r, cache=TRUE, echo = TRUE}
saveRDS(fit.matched, file = "data/match.RDS")   
```

### TMLE in RHC data

There are other papers that have used RHC data [@keele2021comparing;@keele2018pre]. 

```{block, type='rmdcomment'}
@keele2021comparing used TMLE (with super learner SL) method in estimating the impact of RHC on length of stay, and found point estimate $2.01 (95\% CI: 0.6-3.41)$.  
```

In today's workshop, we will learn about this **TMLE-SL** methods.


