# G-computation using ML

```{r setup01, include=FALSE}
require(knitr)
require(glmnet)
require(kableExtra)
require(dplyr)
require(xgboost)
require(SuperLearner)
options(knitr.kable.NA = '')
cachex=TRUE
```

## G-comp using Regression tree

```{r reg2sl, cache=cachex, echo = TRUE}
# Read the data saved at the last chapter
ObsData <- readRDS(file = "data/rhcAnalytic.RDS")
dim(ObsData)
```

### A tree based algorithm

XGBoost is a fast version of gradient boosting algorithm. Let us use this one to fit the data first.

```{r ML1, cache=cachex, echo = TRUE}
require(xgboost)
Y <-ObsData$Y
ObsData.matrix <- model.matrix(out.formula, data = ObsData)
fit3 <- xgboost(data = ObsData.matrix, 
                label = Y,
                max.depth = 10, 
                eta = 1, 
                nthread = 15, 
                nrounds = 100, 
                alpha = 0.5,
                objective = "reg:squarederror", 
                verbose = 0)
```


### Extract outcome prediction as if everyone is treated

```{r ML12, cache=cachex, echo = TRUE}
ObsData.matrix.A1 <- ObsData.matrix 
ObsData.matrix.A1[,"A"] <- 1
ObsData$Pred.Y1 <- predict(fit3, newdata = ObsData.matrix.A1)
summary(ObsData$Pred.Y1)
```

### Extract outcome prediction as if everyone is untreated


```{r ML13, cache=cachex, echo = TRUE}
ObsData.matrix.A0 <- ObsData.matrix
ObsData.matrix.A0[,"A"] <- 0
ObsData$Pred.Y0 <- predict(fit3, newdata = ObsData.matrix.A0)
summary(ObsData$Pred.Y0)
```

### Treatment effect estimate


```{r ML13b, cache=cachex, echo = TRUE}
ObsData$Pred.TE <- ObsData$Pred.Y1 - ObsData$Pred.Y0  

```


Mean value of predicted treatment effect 

```{r reg2acnx1b, cachex=TRUE, echo = TRUE}
summary(ObsData$Pred.TE)
```

## G-comp using regularized methods

### A regularized model

LASSO is a regularized method. Let us use this one to fit the data first.

```{r ML1r, cache=cachex, echo = TRUE}
require(glmnet)
Y <-ObsData$Y
ObsData.matrix <- model.matrix(out.formula, data = ObsData)
fit4 <-  cv.glmnet(x = ObsData.matrix, 
                y = Y,
                alpha = 0.5,
                relax=TRUE)
```


### Extract outcome prediction as if everyone is treated

```{r ML12r, cache=cachex, echo = TRUE}
ObsData.matrix.A1 <- ObsData.matrix 
ObsData.matrix.A1[,"A"] <- 1
ObsData$Pred.Y1 <- predict(fit4, newx = ObsData.matrix.A1,
                           s = "lambda.min")
summary(ObsData$Pred.Y1)
```

### Extract outcome prediction as if everyone is untreated


```{r ML13r, cache=cachex, echo = TRUE}
ObsData.matrix.A0 <- ObsData.matrix
ObsData.matrix.A0[,"A"] <- 0
ObsData$Pred.Y0 <- predict(fit4, newx = ObsData.matrix.A0,
                           s = "lambda.min")
summary(ObsData$Pred.Y0)
```

### Treatment effect estimate


```{r ML13br, cache=cachex, echo = TRUE}
ObsData$Pred.TE <- ObsData$Pred.Y1 - ObsData$Pred.Y0  

```


Mean value of predicted treatment effect 

```{r reg2acnx1br, cachex=TRUE, echo = TRUE}
summary(ObsData$Pred.TE)
```


## G-comp using SuperLearner

### SuperLearner

SuperLearner is an ensemble learning method. Let us use this one to fit the data first.

```{r ML1s, cache=cachex, echo = TRUE, results='hide', warning = FALSE}
require(SuperLearner)
ObsData.noY <- dplyr::select(ObsData, !Y)
fit.sl <- SuperLearner(Y=ObsData$Y, 
                       X=ObsData.noY, 
                       cvControl = list(V = 3),
                       SL.library=c("SL.glm", "SL.glmnet", "SL.xgboost"), 
                       family="gaussian")
```

```{r ML12stestrisk, cache=cachex, echo = TRUE}
fit.sl$cvRisk
```


```{r ML12stestcoef, cache=cachex, echo = TRUE}
fit.sl$coef
```


```{r ML12stest, cache=cachex, echo = TRUE}
all.pred <- predict(fit.sl, type = "response")
head(all.pred$library.predict)
round(all.pred$pred[1:6,],1)
```


### Extract outcome prediction as if everyone is treated

```{r ML12s, cache=cachex, echo = TRUE}
ObsData.noY$A <- 1
ObsData$Pred.Y1 <- predict(fit.sl, newdata = ObsData.noY,
                           type = "response")$pred
summary(ObsData$Pred.Y1)
```

### Extract outcome prediction as if everyone is untreated


```{r ML13s, cache=cachex, echo = TRUE}
ObsData.noY$A <- 0
ObsData$Pred.Y0 <- predict(fit.sl, newdata = ObsData.noY,
                           type = "response")$pred
summary(ObsData$Pred.Y0)
```

### Treatment effect estimate


```{r ML13bs, cache=cachex, echo = TRUE}
ObsData$Pred.TE <- ObsData$Pred.Y1 - ObsData$Pred.Y0  

```


Mean value of predicted treatment effect 

```{r reg2acnx1bs, cachex=TRUE, echo = TRUE}
summary(ObsData$Pred.TE)
```


