# IPTW

```{block, type='rmdcomment'}
In this chapter, we are primarily interested about **exposure modelling** (e.g., fixing imbalance first, before doing outcome analysis).
```

```{r setup01i, include=FALSE}
require(knitr)
require(glmnet)
require(kableExtra)
require(dplyr)
require(xgboost)
require(SuperLearner)
require(Publish)
require(tableone)
require(survey)
require(cobalt)
require(WeightIt)
options(knitr.kable.NA = '')
cachex=TRUE
```

```{r reg2ps, cache=cachex, echo = TRUE}
# Read the data saved at the last chapter
ObsData <- readRDS(file = "data/rhcAnalytic.RDS")
baselinevars <- names(dplyr::select(ObsData, !c(A,Y)))
```

## IPTW steps 

**Modelling Steps**:

According to @austin2011tutorial, we need to follow 4 steps:

|  |  |
|-|-|
|Step 1| exposure modelling: $PS = Prob(A=1|L)$|
|Step 2| Convert $PS$ to $IPW$  = $\frac{A}{PS} + \frac{1-A}{1-PS}$|
|Step 3| Assess balance in weighted sample ($PS$ and $L$)|
|Step 4| outcome modelling: $E(Y|A=1)$ to obtain treatment effect estimate  |

## Step 1: exposure modelling

```{block, type='rmdcomment'}
Exposure modelling: $PS = Prob(A=1|L)$
```

```{r ps1, cache=cachex, echo = TRUE}
ps.formula <- as.formula(paste("A ~",
                               paste(baselinevars,
                                     collapse = "+")))
ps.formula
```

- Other than main effect terms, what other model specifications are possible?
  - Common terms to add (indeed based on biological plausibility; requiring subject area knowledge)
  - Interactions
  - polynomials or splines
  - transformations

Fit logistic regression to estimate propensity scores

```{r ps321xxx, cache=TRUE, echo = TRUE}
PS.fit <- glm(ps.formula,family="binomial", data=ObsData)
require(Publish)
publish(PS.fit,  format = "[u;l]")
```

- Coef of PS model fit is not of concern 
- Model can be rich: to the extent that prediction is better
- But look for multi-collinearity issues
  - SE too high?

Obtain the propesnity score (PS) values from the fit

```{r psx2, cache=TRUE, echo = TRUE}
ObsData$PS <- predict(PS.fit, type="response")
```

Check summaries: 

- enough overlap?
- PS values very close to 0 or 1?

```{r psx2b, cache=TRUE, echo = TRUE}
summary(ObsData$PS)
tapply(ObsData$PS, ObsData$A, summary)
plot(density(ObsData$PS[ObsData$A==0]), 
     col = "red", main = "")
lines(density(ObsData$PS[ObsData$A==1]), 
      col = "blue", lty = 2)
legend("topright", c("No RHC","RHC"), 
       col = c("red", "blue"), lty=1:2)
```


## Step 2: Convert PS to IPW 

```{block, type='rmdcomment'}
Convert $PS$ to $IPW$  = $\frac{A}{PS} + \frac{1-A}{1-PS}$
```

- Convert PS to IPW using the formula. We are using the formula for average treatment effect (ATE). 
- It is possible to use alternative formulas, but we are using ATE formula for our illustration.

```{r psx2c, cache=TRUE, echo = TRUE}
ObsData$IPW <- ObsData$A/ObsData$PS + (1-ObsData$A)/(1-ObsData$PS)
summary(ObsData$IPW)
```

Also possible to use pre-packged software packages to do the same:

```{r psx2c2, cache=TRUE, echo = TRUE}
require(WeightIt)
W.out <- weightit(ps.formula, 
                    data = ObsData, 
                    estimand = "ATE",
                    method = "ps")
summary(W.out$weights)
```


## Step 3: Balance checking

```{block, type='rmdcomment'}
Assess balance in weighted sample ($PS$ and $L$)
```

We can check balance numerically. 

- We set SMD = 0.1 as threshold for balance.
- $SMD \gt 0.1$ means we do not have balance

```{r balp, cache=TRUE, echo = TRUE, fig.height=10, fig.width=5}
require(cobalt)
bal.tab(W.out, un = TRUE, 
        thresholds = c(m = .1))
```

- We can also check this in a plot

```{r balp2, cache=TRUE, echo = TRUE, fig.height=10, fig.width=5}
require(cobalt)
love.plot(W.out, binary = "std",
          thresholds = c(m = .1),
          abs = TRUE, 
          var.order = "unadjusted", 
          line = TRUE)
```

```{block, type='rmdcomment'}
All covariates are balanced! Reverse engineered an RCT!?!
```

## Step 4: outcome modelling 

```{block, type='rmdcomment'}
Outcome modelling: $E(Y|A=1)$ to obtain treatment effect estimate
```

Estimate the effect of treatment on outcomes 

```{r psx4, cache=TRUE, echo = TRUE}
out.formula <- as.formula(Y ~ A)
out.fit <- glm(out.formula,
               data = ObsData,
               weights = IPW)
publish(out.fit)
```


```{r, cache=TRUE, echo = TRUE}
saveRDS(out.fit, file = "data/ipw.RDS")
```
